---
title: "Group_1_Analysis"
author: "Ying Yang, Linbin Lai"
number-sections: true
format:
  html:
    embed-resources: true
    code-tools: true
  pdf: 
    keep-tex: true
    include-in-header: 
       text: |
         \usepackage{fvextra}
         \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
         \DefineVerbatimEnvironment{OutputCode}{Verbatim}{breaklines,commandchars=\\\{\}}
editor_options:
  chunk_output_type: console
execute:
  echo: true
  eval: true
  warning: false
  message: false
---

## Forward Selection

### Load Package and Document

```{r}
rm(list = ls())
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(MASS)
library(dplyr)
library(janitor)
library(tidyr)
FIES <- read.csv("~/Data Analysis Skills/Group project2/dataset01.csv")
unique(FIES$Region)#ignore the region column
FIES <- FIES %>%
  dplyr::select(Total.Number.of.Family.members, Total.Household.Income, Total.Food.Expenditure, Household.Head.Sex, Household.Head.Age, Type.of.Household, House.Floor.Area, House.Age, Number.of.bedrooms, Electricity)
FIES$Total.Number.of.Family.members <- as.factor(FIES$Total.Number.of.Family.members)
FIES$Household.Head.Sex <- as.factor(FIES$Household.Head.Sex)
FIES$Type.of.Household <- as.factor(FIES$Type.of.Household)
FIES$Electricity <- as.factor(FIES$Electricity)
levels(FIES$Electricity) <- c("No", "Yes")
FIES$Number.of.bedrooms <- as.factor(FIES$Number.of.bedrooms)
levels(FIES$Number.of.bedrooms) <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9")
```

### Analyze each element individually

#### `Total.Number.of.Family.members` vs `Total.Household.Income`

```{r}
#| label: fig-boxplot1
#| fig-subcap: Annual household income by Number of family members.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, y = Total.Household.Income, fill = Total.Number.of.Family.members)) + 
  geom_boxplot() + 
  labs(x = "Number of people living in the house.", y = "Annual household income (in Philippine peso)") + 
  theme(legend.position = "none")
```

From the boxplot, we can find that median annual income appears to be similar for most household sizes (number of people), particularly from one to six person households. Moreover, as household size increases, the number of households with unusually high incomes decreases.

#### `Total.Number.of.Family.members` vs `Total.Food.Expenditure`

```{r}
#| label: fig-boxplot2
#| fig-subcap: Annual expenditure by the household on food by Number of family members.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, y = Total.Household.Income, fill = Total.Number.of.Family.members)) + 
  geom_boxplot() + 
  labs(x = "Number of people living in the house.", y = "Annual expenditure by the household on food") + 
  theme(legend.position = "none")
```

From this boxplot, median food expenditures appear to increase gradually as household size increases, especially from one to nine person households. There are fewer data points for larger households (e.g. more than 10 people), as can be seen from the width of the boxplot, with smaller box widths indicating smaller sample sizes for these household sizes.

#### `Total.Number.of.Family.members` vs `Household.Head.Sex`

```{r}
FIES %>%
  tabyl(Household.Head.Sex, Total.Number.of.Family.members) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns() #To show original counts
```

```{r}
#| label: fig-barplot1
#| fig-subcap: Barplot of Number of people living in the house by Head of the households sex.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, group = Household.Head.Sex)) + 
  geom_bar(aes(y = ..prop.., fill = Household.Head.Sex), stat = "count", position = "dodge") + 
  labs(x = "Number of people living in the house", y = "Proportion")
```

For different family sizes, families headed by men generally account for a larger proportion, especially in families with three to seven people. Among one person, two person and three person households, the proportion of female-headed households is higher than the male-headed households.

#### `Total.Number.of.Family.members` vs `Household.Head.Age`

```{r}
#| label: fig-boxplot3
#| fig-subcap: Head of the households age by Number of family members.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, y = Household.Head.Age, fill = Total.Number.of.Family.members)) + 
  geom_boxplot() + 
  labs(x = "Number of people living in the house.", y = "Head of the households age (in years)") + 
  theme(legend.position = "none")
```

For different family sizes, the median age of the household head is relatively stable among different family sizes, mainly concentrated around 50 years old. So I think there is no obvious correlation trend between the age of the household head and the size of the family.

#### `Total.Number.of.Family.members` vs `Type.of.Household`

```{r}
#| label: fig-barplot2
#| fig-subcap: Barplot of Number of people living in the house by the relationship between the group of people living in the house.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, group = Type.of.Household)) + 
  geom_bar(aes(y = ..prop.., fill = Type.of.Household), stat = "count", position = "dodge") + 
  labs(x = "Number of people living in the house", y = "Proportion")
```

There is a correlation between household type and the number of people living in the house, with extended households being more common in medium and large households, while single households predominate among small households. And for two or More Non-related Persons/Members, it has highly proportion in two person and three person household.

#### `Total.Number.of.Family.members` vs `House.Floor.Area`

```{r}
#| label: fig-boxplot4
#| fig-subcap: Floor area of the house by Number of family members.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, y = House.Floor.Area, fill = Total.Number.of.Family.members)) + 
  geom_boxplot() + 
  labs(x = "Number of people living in the house.", y = "Floor area of the house (in meter square)") +
  theme(legend.position = "none")
```

Median house floor area appears to be relatively stable for most number of people living in the house, without increasing significantly as the number of residents increases. The graph shows a large number of outliers, indicating that some households have homes that are well outside the typical range for similar household sizes.Therefore, house floor area does not appear to increase significantly with number of people living in the house.

#### `Total.Number.of.Family.members` vs `House.Age`

```{r}
#| label: fig-boxplot5
#| fig-subcap: Age of the building by Number of family members.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, y = House.Age, fill = Total.Number.of.Family.members)) + 
  geom_boxplot() + 
  labs(x = "Number of people living in the house.", y = "Age of the building (in years)") +
  scale_fill_brewer(palette = "Set3") +#change to different color
  theme(legend.position = "none")
```

The distribution range of house ages is relatively consistent, suggesting that the distribution of house ages is similar across household sizes. In conclude, there does not appear to be a direct relationship between house age and household size, with the median age remaining relatively stable across household sizes.

#### `Total.Number.of.Family.members` vs `Number.of.bedrooms`

```{r}
#| label: fig-barplot3
#| fig-subcap: Barplot of Number of people living in the house by Number of bedrooms in the house.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, group = Number.of.bedrooms)) + 
  geom_bar(aes(y = ..prop.., fill = Number.of.bedrooms), stat = "count", position = "dodge") + 
  labs(x = "Number of people living in the house", y = "Proportion")
```

As the number of residents increases, the median number of bedrooms gradually increases, but this growth is not linear.

#### `Total.Number.of.Family.members` vs `Electricity`

```{r}
#| label: fig-barplot4
#| fig-subcap: Barplot of Number of people living in the house by the electricity.
ggplot(data = FIES, aes(x = Total.Number.of.Family.members, group = Electricity)) + 
  geom_bar(aes(y = ..prop.., fill = Electricity), stat = "count", position = "dodge") + 
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Number of people living in the house", y = "Proportion")
```

For most of the families, regardless of size, they have electricity. Among small scale households (especially single person households), the proportion of households without electricity is relatively high. As household size increases, the proportion of households without electricity gradually decreases. In larger households (seven people and above), almost all households have electricity supply.

### Summary of Analyze each element individually

Overall, according to the separate analysis of explanatory variables, the variables that should be retained for now are:Annual household income, Annual expenditure by the household on food, Head of the households sex, Relationship between the group of people living in the house and Electricity.

### Analysis of the overall part

```{r}
FIES <- FIES %>%
  dplyr::select(Total.Number.of.Family.members, Total.Household.Income, Total.Food.Expenditure, Household.Head.Sex, Type.of.Household, Number.of.bedrooms, Electricity)

FIES$Total.Number.of.Family.members <- as.numeric(as.character(FIES$Total.Number.of.Family.members))

model.FIES <- glm(Total.Number.of.Family.members ~ Total.Household.Income + Total.Food.Expenditure + Household.Head.Sex + Type.of.Household + Electricity, data = FIES, family = poisson(link = "log"))

model.FIES %>%
  summary()
```

`Type.of.HouseholdTwo or More Nonrelated Persons/Members`: The p value is 0.365302, indicating that the effect of this variable is not significant.\
`Number.of.bedrooms (7 to 9 bedrooms)`: The p-values ​​of these variables range from 0.129907 to 0.371571, indicating that their impact on the number of family members is not significant.\
`Electricity Yes`: The p-value is 0.462128, which means that the presence or absence of electricity has no significant impact on the number of household members.

```{r}
#| label: fig-plot
#| fig-subcap: Rate ratios plot.
plot_model(model.FIES, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.25)
```

From total household income and total food expenditure, their rate ratios is exactly 1.00, they prove no effect on the number of family members. Household Head Sex's rate ratio is 1.28, which suggests that households headed by males have a 28% increase in the expected count of family members compared to households headed by females, assuming all other factors are held constant. Type of Household has the rate ratios about 0.72 indicates that single-family households have a 28% decrease in the expected count of family members compared to the reference household type (likely non-single family households). And for type of household With an rate ratios of 0.87, this suggests that households consisting of two or more non-related persons/members have a 13% decrease in the expected count of family members compared to the reference category. Number of Bedrooms suggesting a small decrease of 3% in the expected count of family members for each additional bedroom in the household.

### Result

At last , we keep `Household.Head.Sex` as explanatory variables: $$
Total.Number.of.Family.members = \beta_0 + \beta_1 \times \text{Household.Head.Sex}
$$

```{r}
FIES$Total.Number.of.Family.members <- as.numeric(as.character(FIES$Total.Number.of.Family.members))

model.FIES <- glm(Total.Number.of.Family.members ~ Household.Head.Sex, data = FIES, family = poisson(link = "log"))

model.FIES %>%
  summary()
model <- glm.nb(Total.Number.of.Family.members ~ Household.Head.Sex,data = FIES)
summary(model)
model$aic
```

## Backward Elimination

### Find the suitable generalized linear model using family=poisson(Backward regression)

```{r}
data <- read.csv('~/Data Analysis Skills/Group project2/dataset01.csv')
house_data1 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age+
                  Number.of.bedrooms+
                  Electricity,
                family = poisson, data=data)
house_data1%>%
  summary()
```

The variable Electricity has p-value greater than 0.05, so delete this variable and refit the model.

```{r}
house_data2 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age+
                  Number.of.bedrooms,
                family = poisson, data=data)
house_data2%>%
  summary()
```

The variable Number.of.bedrooms has p-value greater than 0.05, so delete this variable and refit the model.

```{r}
house_data3 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age,
                family = poisson, data=data)
house_data3%>%
  summary()
```

The variable Type.of.Household has p-value greater than 0.05, so delete this variable and refit the model.

```{r}
house_data4 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age,
                family = poisson, data=data)
house_data4%>%
  summary()
```

The variable Household.Head.Age has p-value greater than 0.05, so delete this variable and refit the model.

```{r}
house_data5 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Total.Number.of.Family.members+
                  House.Floor.Area+
                  House.Age,
                family = poisson, data=data)
house_data5%>%
  summary()
```

The variable House.Floor.Area has p-value greater than 0.05, so delete this variable and refit the model.

```{r}
house_data6 <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Total.Number.of.Family.members+
                  House.Age,
                family = poisson, data=data)
house_data6%>%
  summary()
```

The model Total.Number.of.Family.members\~Total.Household.Income+Total.Food.Expenditure+Household.Head.Sex+Household.Head.Age+Type.of.Household+Total.Number.of.Family.members+House.Floor.Area+House.Age+Number.of.bedrooms has the lowest AIC which is 7014.1.

After reducing variables which with p-value large than 0.05, the model becomes as follows:

Total.Number.of.Family.members\~Total.Household.Income+Total.Food.Expenditure+Household.Head.Sex+Household.Head.Age+Total.Number.of.Family.members+House.Age

Therefore, these five variables(Total.Household.Income, Total.Food.Expenditure, Household.Head.Sex, Household.Head.Age and House.Age) are significant. So the household related variables(Total.Household.Income, Total.Food.Expenditure, Household.Head.Sex, Household.Head.Age and House.Age)influence the number of people living in a household.

We see that the coefficient for Total.Household.Income is negative, indicating that if a household has more total household income, there will be less number of people living in the household. Similarly, the coefficient for Total.Food.Expenditure is positive, which means as the total food expenditure increase, there will be more people living in the household. The Household.Head.SexMale coefficient is positive, indicating a higher number of people living in a household for male house head. Besides, the coefficient for Household.Head.Age is negative, that is, as household head age increases there will be less member living in the house. Finally, the House.Age coefficient is negative, which means if the house get older, there will be less number of people living in the household.

#### Odds ratios

```{r}
#| label: fig-oddplot
#| fig-subcap: The odds ratios plot.
plot_model(house_data2, show.values = TRUE,
title = "", show.p = FALSE, value.offset = 0.25)
```

We interpret the odds ratios as follows: women’s odds of number of family members were 0.8 times those of man. For each unit change for Total.Household.Income, Total.Food.Expenditure,Household.Head.Age and House.Age the odds of number of family members do not change, which means they prove no effect on the number of family members.

Therefore, the best model is:

$$Total.Number.of.Family.members=\beta_0 + \beta_1* \cdot \mathbb{I}_{\text{Household.Head.SexFemale}}$$ $$=(1.533e+00)-(2.212e-01)\cdot \mathbb{I}_{\text{Household.Head.SexFemale}}$$

$\mathbb{I}_i$ is is an indicator function such that representing each gender such that $$\mathbb{I}_i=\left\{
\begin{array}{ll}
1 ~~~ \mbox{if the household head is female},\\
0 ~~~ \mbox{Otherwise}.\\
\end{array}
\right.$$

### Data visualization: Analysis for significant explanatory variable(Household.Head.Sex) with response variable(Total.Number.of.Family.members)

#### Household.Head.Sex\~Total.Number.of.Family.members

```{r}
data %>%
  tabyl(Household.Head.Sex, Total.Number.of.Family.members) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns()
```

```{r}
ggplot(data = data, aes(x = Total.Number.of.Family.members, group = Household.Head.Sex)) +
  geom_bar(aes(y = after_stat(prop), fill = Household.Head.Sex), stat = "count", position = "dodge") +
  labs(x = "Total.Number.of.Family.members", y = "Proportion")
```

There is a clear pattern that female has higher proportion in smaller family size than male.

### Find the suitable Negative Binomial generalized linear model(Backward regression)

```{r}
library(MASS)
data$Total.Number.of.Family.members <- as.numeric(as.character(data$Total.Number.of.Family.members))
housenb_data1 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age+
                  Number.of.bedrooms+
                  Electricity,
                  data=data)
housenb_data1%>%
  summary()
```

The variable Electricity has big p-value, so we reduce this variable and refit the model.

```{r}
housenb_data2 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age+
                  Number.of.bedrooms,
                  data=data)
housenb_data2%>%
  summary()
```

The variable Number.of.bedrooms has big p-value, so we reduce this variable and refit the model.

```{r}
housenb_data3 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+
                  House.Age,
                  data=data)
housenb_data3%>%
  summary()
```

The variable Type.of.Household has big p-value, so we reduce this variable and refit the model.

```{r}
housenb_data4 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Total.Number.of.Family.members+
                  House.Floor.Area+
                  House.Age,
                  data=data)
housenb_data4%>%
  summary()
```

The variable Household.Head.Age has big p-value, so we reduce this variable and refit the model.

```{r}
housenb_data5 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Total.Number.of.Family.members+
                  House.Floor.Area+
                  House.Age,
                  data=data)
housenb_data5%>%
  summary()
```

The variable House.Floor.Area has big p-value, so we reduce this variable and refit the model.

```{r}
housenb_data6 <- glm.nb(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Total.Number.of.Family.members+
                  House.Age,
                  data=data)
housenb_data6%>%
  summary()
```

#### Odds ratios

```{r}
plot_model(housenb_data2, show.values = TRUE,
title = "", show.p = FALSE, value.offset = 0.25)
```

So the negative binomial regression method's result is the same as poisson.
