---
title: "Group_1_Analysis"
author: "Ying Yang"
number-sections: true
format:
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options:
  chunk_output_type: inline
execute:
  echo: true
  eval: true
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(dplyr)
library(janitor)
```

# Find the suitable glm model

## using family=poisson

```{r}
data <- read.csv('dataset01.csv')
house_data <- glm(Total.Number.of.Family.members~
                  Total.Household.Income+
                  Total.Food.Expenditure+
                  Household.Head.Sex+
                  Household.Head.Age+
                  Type.of.Household+
                  Total.Number.of.Family.members+
                  House.Floor.Area+House.Age+
                  Number.of.bedrooms+
                  Electricity,
                family = poisson, data=data)
house_data%>%
  summary()
```

The AIC for the model is very large, so we do not consider model which family=poisson

## using family=binomial

```{r}
data <- read.csv('dataset01.csv')
data$Household.Head.Sex <- as.factor(data$Household.Head.Sex)
data$Type.of.Household <- as.factor(data$Type.of.Household)
data$Total.Number.of.Family.members <- as.factor(data$Total.Number.of.Family.members)
levels(data$Total.Number.of.Family.members) <- c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15")
data$Number.of.bedrooms <- as.factor(data$Number.of.bedrooms)
data$Electricity <- as.factor(data$Electricity)
```

```{r}
mod.data1 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Household.Head.Age+Type.of.Household+Total.Number.of.Family.members+House.Floor.Area+House.Age+Number.of.bedrooms+Electricity,data=data,
                family = binomial(link = "logit"))
mod.data1%>%
  summary()
```

The variable Type.of.Household has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data2 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Household.Head.Age+Total.Number.of.Family.members+House.Floor.Area+House.Age+Number.of.bedrooms+Electricity,data=data,
                family = binomial(link = "logit"))
mod.data2%>%
  summary()
```

The variable Household.Head.Age has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data3 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Total.Number.of.Family.members+House.Floor.Area+House.Age+Number.of.bedrooms+Electricity,data=data,
                family = binomial(link = "logit"))
mod.data3%>%
  summary()
```

The variable House.Age has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data4 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Total.Number.of.Family.members+House.Floor.Area+Number.of.bedrooms+Electricity,data=data,
                family = binomial(link = "logit"))
mod.data4%>%
  summary()
```

The variable Number.of.bedrooms has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data5 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Total.Number.of.Family.members+House.Floor.Area+Electricity,data=data,
                family = binomial(link = "logit"))
mod.data5%>%
  summary()
```

The variable Electricity has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data6 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Total.Number.of.Family.members+House.Floor.Area,data=data,
                family = binomial(link = "logit"))
mod.data6%>%
  summary()
```

The variable Electricity has p-value greater than 0.05, so delate this variable and refit the model.

```{r}
mod.data7 <- glm(Total.Number.of.Family.members~Total.Household.Income++Total.Food.Expenditure+Household.Head.Sex+Total.Number.of.Family.members,data=data,
                family = binomial(link = "logit"))
mod.data7%>%
  summary()
```

All variables' p-values are less than 0.05, which means these three variables(Total.Household.Income, Total.Food.Expenditure and Household.Head.Sex) are significant. So the household related variables(Total.Household.Income, Total.Food.Expenditure and Household.Head.Sex)influence the number of people living in a household.

We see that the coefficient for Total.Household.Income is negative, indicating that if a household has more total household income, there will be less number of people living in the household. Similarly, the coefficient for Total.Food.Expenditure is positive, which means as the total food expenditure increase, there will be more people living in the household. Finally the Household.Head.SexMale coefficient is positive, indicating a higher number of people living in a household for male house head.

### Odds ratios

```{r}
plot_model(mod.data, show.values = TRUE,
title = "", show.p = FALSE, value.offset = 0.25)
```

We interpret the odds ratios as follows: men’s odds of number of family members were 2.25 times those of women. For each unit change for Total.Household.Income and Total.Food.Expenditure the odds of number of family members do not change.

### Data visulization: Analysis for each significant explanatory variable with response variable(Total.Number.of.Family.members)

#### 1. Total.Household.Income\~Total.Number.of.Family.members

```{r}
ggplot(data = data, aes(x = Total.Number.of.Family.members, y = Total.Household.Income, fill = Total.Number.of.Family.members)) +
  geom_boxplot() +
  labs(x = "Total.Number.of.Family.members", y = "Total.Household.Income") +
  theme(legend.position = "none")
```

For households which has more members, will have more total household income.

#### 2. Total.Food.Expenditure\~Total.Number.of.Family.members

```{r}
ggplot(data = data, aes(x = Total.Number.of.Family.members, y = Total.Food.Expenditure, fill = Total.Number.of.Family.members)) +
  geom_boxplot() +
  labs(x = "Total.Number.of.Family.members", y = "Total.Food.Expenditure") +
  theme(legend.position = "none")
```

For households which has more members, will have more total food expenditure.

#### 3. Household.Head.Sex\~Total.Number.of.Family.members

```{r}
data %>%
  tabyl(Household.Head.Sex, Total.Number.of.Family.members) %>%
  adorn_percentages() %>%
  adorn_pct_formatting() %>%
  adorn_ns() # To show original counts
```

```{r}
ggplot(data = data, aes(x = Total.Number.of.Family.members, group = Household.Head.Sex)) +
  geom_bar(aes(y = after_stat(prop), fill = Household.Head.Sex), stat = "count", position = "dodge") +
  labs(x = "Total.Number.of.Family.members", y = "Proportion")
```

There is a clear pattern that female has higher proportion in smaller family size than male.
